### 2007-01-11 Documentation/part1.tex
JK   - Documented new functions which return pointers to arrays of bin widths and     bin central energies   - Adjusted documentation of tilt   - Added warning in the section about energy calibration errors, saying that it     will not work with variable bin widths.

In some cases, it is necessary to obtain information from the \AEDL\
files. Very basic information is supplied with
\begin{function}
\GLBNS{GetEminEmax}
{\tt int glbGetEminEmax(int experiment, double *emin, double *emax)} returns the
energy range [{\tt emin},{\tt emax}] covered by the binning as defined in the AEDL file for the experiment
{\tt experiment}.
\end{function}
\begin{function}
\GLBNS{GetEnergyWindow}
{\tt int glbGetEnergyWindow(int exp, int rule, double *low, double *high)} returns the
energy window [{\tt low},{\tt high}] defined in the AEDL file for the experiment
{\tt exp} and rule {\tt rule}.
\end{function}
\begin{function}
\GLBNS{GetEnergyWindowBins}
{\tt int glbGetEnergyWindowBins(int exp, int rule, int *lowbin, int *highbin)} returns the
energy window in terms of bin numbers [{\tt lowbin},{\tt highbin}] corresponding to the energy window defined in the AEDL file for the experiment {\tt exp} and rule {\tt rule}.
\end{function}
\begin{function}
\GLBNS{GetNumberOfBins}
{\tt int glbGetNumberOfBins(int exp)} returns the number of bins for the experiment {\tt exp}.
\end{function}
\begin{function}
\GLBNS{GetNumberOfSamplingPoints}
{\tt int glbGetNumberOfSamplingPoints(int exp)} returns the number of sampling points for the experiment {\tt exp}.
\end{function}

>>>>

In some cases, it is necessary to obtain information from the \AEDL\
files. Very basic information is supplied with
\begin{function}
\GLBNS{GetEminEmax}
{\tt int glbGetEminEmax(int experiment, double *emin, double *emax)} returns the
energy range [{\tt emin},{\tt emax}] covered by the binning as defined in the AEDL file for the experiment
{\tt experiment}.
\end{function}
\begin{function}
\GLBNS{GetEnergyWindow}
{\tt int glbGetEnergyWindow(int exp, int rule, double *low, double *high)} returns the
energy window [{\tt low},{\tt high}] defined in the AEDL file for the experiment
{\tt exp} and rule {\tt rule}.
\end{function}
\begin{function}
\GLBNS{GetEnergyWindowBins}
{\tt int glbGetEnergyWindowBins(int exp, int rule, int *lowbin, int *highbin)} returns the
energy window in terms of bin numbers [{\tt lowbin},{\tt highbin}] corresponding to the energy window defined in the AEDL file for the experiment {\tt exp} and rule {\tt rule}.
\end{function}
\begin{function}
\GLBNS{GetNumberOfBins}
{\tt int glbGetNumberOfBins(int exp)} returns the number of bins for the experiment {\tt exp}.
\end{function}
\begin{function}
\GLBNS{GetNumberOfSamplingPoints}
{\tt int glbGetNumberOfSamplingPoints(int exp)} returns the number of sampling points for the experiment {\tt exp}.
\end{function}
\begin{function}
\GLBNS{GetBinSizeListPtr}
{\tt double *glbGetBinSizeListPtr(int exp)} returns a pointer to the array of bin widths
for the experiment {\tt exp}.
\end{function}
\begin{function}
\GLBNS{GetSamplingStepsizeListPtr}
{\tt double *glbGetSamplingStepsizeListPtr(int exp)} returns a pointer to the array of
sampling step sizes for the experiment {\tt exp}.
\end{function}
\begin{function}
\GLBNS{GetBinCentersListPtr}
{\tt double *glbGetBinCentersListPtr(int exp)} returns a pointer to the array of
mean bin energies for the experiment {\tt exp}.
\end{function}
\begin{function}
\GLBNS{GetSamplingPointsListPtr}
{\tt double *glbGetSamplingPointsListPtr(int exp)} returns a pointer to the array of
sampling points for the experiment {\tt exp}.
\end{function}

### 2007-01-11 Documentation/part2.tex
JK   - Documented new functions which return pointers to arrays of bin widths and     bin central energies   - Adjusted documentation of tilt   - Added warning in the section about energy calibration errors, saying that it     will not work with variable bin widths.

For the parameterization of an energy calibration error, two possibilities
are implemented. The first one (method ``T'') is somewhat simpler, 
whereas the second one (method ``C'')
is more accurate, but it requires a careful choice of parameters. 
The first option (method ``T'') is\footnote{Note that this behaviour
has slightly changed compared to previous \GLOBES\ releases.}
\begin{equation}
s_i(a,b) \equiv s_i(a)+b\cdot s_i \cdot i / N_\mathrm{bins}.
\end{equation}
It is often refered to as a ``tilt'' of the
spectrum, since it describes a linear distortion 
of the event rate spectrum. Note that this definition of a tilt makes
only sense in combination with a large enough normalization error, since
the tilt may also affect the normalization. 
%
The second option (method ``C'') is closer to an actual energy
calibration error, which means that one should test this option whenever
one suspects a large impact of this systematical error.
It is based upon replacing the events in the $i$th bin (mean reconstructed energy $E'_i$)
by the ones at the energy $(1+b)\cdot E'_i$. If the target energy does not exactly hit
a (discrete) bin energy $E_k$, linear interpolation is used. We use the following approximation:
\begin{eqnarray}
s_i(a,b)&=& (1+b)\cdot \left[ \left( s_{k+1}(a)-s_k(a) \right)\cdot (\delta-k) +s_k(a) \right] \,,\\
\delta&=&b\cdot(i+t_0+ 1/2)+i\,,\nonumber\\
k&=& \mathrm{div}(\delta,1)\,,\nonumber\\
t_0&=&E'_\mathrm{min}/\Delta E_0\,.\nonumber
\end{eqnarray}
Here, $E'_\mathrm{min} = {\tt \$emin}$,  $\Delta E_0$ is the bin width
({\tt \$emax}-{\tt \$emin})/{\tt \$bins},
and ``div'' refers to the integer part of the division. Note that the
factor $(1+b)$ in the first equation comes from a renormalization of
the bin width, since also the bin width is altered by the replacement of
the energies. Furthermore, special care has to be payed to the limits
 $k<1$ or $k+1>N_\mathrm{bins}$, since there $s_k$ or $s_{k+1}$ may not have been calculated. By default, it is assumed that $s_k$ is
zero in those cases. However, if the event rates are still large at
the limits, errors will be introduced leading to a wrong
estimate of the impact of the calibration error. In this case,
one should truncate the analysis range by a few bins at the boundaries
and therefore ensure in this way that only those $s_i$ are used whose index
$k$ is within the range $0,\ldots, N_\mathrm{bins}-1$ (\cf, \figu{energies}). 
Thus, it is possible to constrain the analysis energy range 
with each rule to an energy window\index{norm}{Energy!window}:
\index{aedl}{rule@{\tt rule}!energy window@{\tt "@energy\_window}}
\begin{quote}
{\tt 
\tb @energy\_window = 4.0 : 50.0 
}
\end{quote}
The default energy window is given between the minimal and maximal reconstructed energy. To be on the save side, reduce analysis window
compared to the bin range on each side by about three times the
energy calibration error.

>>>>

For the parameterization of an energy calibration error, two possibilities
are implemented. The first one (method ``T'') is somewhat simpler, 
whereas the second one (method ``C'')
is more accurate, but it requires a careful choice of parameters. 
The first option (method ``T'') is\footnote{Note that this behaviour
has slightly changed compared to previous \GLOBES\ releases.}
\begin{equation}
  s_i(a,b) \equiv s_i(a)+b\cdot s_i \, (E'_i - \bar{E'}) / (E'_\mathrm{max}-E'_\mathrm{min}),
\end{equation}
where $E'_\mathrm{min}$ and $E'_\mathrm{max}$ correspond to {\tt \$emin} and
{\tt \$emax}, $\bar{E'} = \tfrac{1}{2} (E'_\mathrm{max}+E'_\mathrm{min})$ is
the median of this energy interval, and $E'_i$ is the mean (reconstructed)
energy of the $i$th bin.
This method is often refered to as a ``tilt'' of the
spectrum, since it describes a linear distortion 
of the event rate spectrum.
%Note that this definition of a tilt makes
%only sense in combination with a large enough normalization error, since
%the tilt may also affect the normalization. 
%
The second option (method ``C'') is closer to an actual energy
calibration error, which means that one should test this option whenever
one suspects a large impact of this systematical error.
It is based upon replacing the events in the $i$th bin
by the ones at the energy $(1+b)\cdot E'_i$. If the target energy does not exactly hit
a (discrete) bin energy $E_k$, linear interpolation is used. We use the following approximation:
\begin{eqnarray}
s_i(a,b)&=& (1+b)\cdot \left[ \left( s_{k+1}(a)-s_k(a) \right)\cdot (\delta-k) +s_k(a) \right] \,,
  \label{eq:Ecalib} \\
\delta&=&b\cdot(i+t_0+ 1/2)+i\,,\nonumber\\
k&=& \mathrm{div}(\delta,1)\,,\nonumber\\
t_0&=&E'_\mathrm{min}/\Delta E_0\,.\nonumber
\end{eqnarray}
Here, $\Delta E_0$ is the bin width ({\tt \$emax}-{\tt \$emin})/{\tt \$bins},
and ``div'' refers to the integer part of the division. It is important
to keep in mind that this definition of the energy calibration error 
makes sense only for constant bin width, so the corresponding $\chi^2$
functions should not be used in conjunction with the {\tt \$binwidth} directive.
The factor $(1+b)$ in \eq~\eqref{eq:Ecalib} comes from a renormalization of
the bin width, since also the bin width is altered by the replacement of
the energies. Furthermore, special care has to be payed to the limits
 $k<1$ or $k+1>N_\mathrm{bins}$, since there $s_k$ or $s_{k+1}$ may not have been calculated. By default, it is assumed that $s_k$ is
zero in those cases. However, if the event rates are still large at
the limits, errors will be introduced leading to a wrong
estimate of the impact of the calibration error. In this case,
one should truncate the analysis range by a few bins at the boundaries
and therefore ensure in this way that only those $s_i$ are used whose index
$k$ is within the range $0,\ldots, N_\mathrm{bins}-1$ (\cf, \figu{energies}). 
Thus, it is possible to constrain the analysis energy range 
with each rule to an energy window\index{norm}{Energy!window}:
\index{aedl}{rule@{\tt rule}!energy window@{\tt "@energy\_window}}
\begin{quote}
{\tt 
\tb @energy\_window = 4.0 : 50.0 
}
\end{quote}
The default energy window is given between the minimal and maximal reconstructed energy. To be on the save side, reduce analysis window
compared to the bin range on each side by about three times the
energy calibration error.
